Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-05-28 paper_enahpc14/Conclusion.tex
updates a few things according to the reviewers' suggestions

We  have   presented  a  methodology  for   comparing  performance  of
\cosmoart,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandybridge (2.6 GHz) system has
lower  time-to-solution but  higher energy-to-solution  than  an Intel
Ivybridge  (2.2  GHz) system,  although  the  two  metrics are  indeed
strongly correlated.  On the  other hand, the usage of energy-friendly
MPI waiting  techniques analysed with  our power-performance framework
over  a  series  of runs  of  \cosmoart  on  \tinto and  power  traces
demonstrate that it  is possible to reduce both  power dissipation and
energy-to-solution   while  maintaining   (or  even   decreasing)  the
time-to-solution. The resulting profiles indicate that simple changes,
such as making use of the  blocking MPI policy rather than the polling
is  possible to  slightly  reduce both  power  dissipation and  energy
consumption (by 2\,\%).

>>>>

We  have   presented  a  methodology  for   comparing  performance  of
\cosmoart,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandybridge (2.6 GHz) system has
lower time-to-solution but higher energy-to-solution than an Intel Ivy
Bridge (2.2 GHz) system, although  the two metrics are indeed strongly
correlated.   On the  other  hand, the  usage  of energy-friendly  MPI
waiting techniques analysed  with our power-performance framework over
a series of  runs of \cosmoart on \tinto  and power traces demonstrate
that   it  is   possible  to   reduce  both   power   dissipation  and
energy-to-solution   while  maintaining   (or  even   decreasing)  the
time-to-solution. The resulting profiles indicate that simple changes,
such as making use of the  blocking MPI policy rather than the polling
is  possible to  slightly  reduce both  power  dissipation and  energy
consumption (by 2\,\%).

### 2014-05-28 paper_enahpc14/Sec2:COSMO-ART.tex
updates a few things according to the reviewers' suggestions

The  model  system  \cosmoart  described in  \cite{Vogel-2009},  is  a
regional   to  continental   scale   model  coupled   online  to   the
\textsc{Cosmo}   numerical  weather   prediction  and   climate  model
\cite{Baldauf-2011}.    It  incorporates  sophisticated   modules  for
gaseous  chemistry   and  aerosol  dynamics  and   allows  the  online
calculation of  reactive trace  substances and their  interaction with
the  atmosphere.    Detailed  model   description  can  be   found  in
\cite{Bangert-2012, Knote-2011, Knote-2013}.

>>>>

The model system \cosmoart (ART stands for Aerosols and Reactive Trace
gases) developed at KIT  Karlsruhe \cite{Vogel-2009}, is a regional to
continental scale model coupled online to the \textsc{Cosmo} numerical
weather  prediction (NWP) and  climate model  \cite{Baldauf-2011}.  It
allows the  online calculation of reactive trace  substances and their
interaction with  the atmosphere.  Physical  processes like transport,
turbulent diffusion,  and dry and wet deposition  are treated together
with photochemistry and aerosol dynamics using the modal approach.

Aerosols dynamics  are simulated  with the modal  aerosol microphysics
module  MADE \cite{Ackermann-1998},  expanded in  MADEsoot  to feature
explicit  treatment of  soot aging  through condensation  of inorganic
\cite{Riemer-2003} and organic  substances. MADE describes the aerosol
population  through  five   modes  representing  sub-micron  particles
consisting of sulphate, ammonium, nitrate, particulate organic matter,
water and  soot \cite{Riemer-2004} in  a range of mixing  state. These
modes are coupled  with the gas phase by  condensation and nucleation,
and are  strongly influenced by  anthropogenic emissions of  gases and
particles. MADEsoot  describes the sub-micrometer  aerosol population,
composed  of  sea salt  \cite{Lundgren-2013},  mineral dust  particles
\cite{Vogel-2006,   Stanelle-2010},  by   means  of   six  interacting
log-normal modes.  For each  mode, mass contributions and total number
concentration are prognostic  quantities, while the standard deviation
is fixed.

Specific  modules are included  to simulate  the dispersion  of pollen
grains    \cite{Vogel-2008}    and    other   biological    particles.
Meteorologically-influenced emissions  are also online  coupled within
the  model  system.  The  biogenic  VOC  (volatile organic  compounds)
emissions are  calculated as functions of  the land use  type based on
the Global  Land Cover 2000  dataset and the modeled  temperatures and
radiative fluxes \cite{Vogel-1995}.

The gaseous chemistry in COSMO-ART  is solved by a modified version of
the  Regional  Acid  Deposition  Model, Version  2  (RADM2)  mechanism
\cite{Stockwell-1990}, which is extended to describe secondary organic
aerosol  formation  based  on  a  volatile basis  set  (VBS)  approach
\cite{Athanasopoulou-2013}  and  hydroxyl  radical  recycling  due  to
isoprene chemistry  \cite{Geiger-2003}.  The thermodynamic equilibrium
between the  gas and particulate  phases of the inorganic  material is
achieved through the ISORROPIA II module \cite{Fountoukis-2007}.

COSMO-ART  is  fully  online-coupled,  and  allows  for  feedbacks  of
aerosols  on  temperature,  radiation  and cloud  condensation  nuclei
(CCN).   Analytical  description  of  modules  incorporated  for  this
purpose  exists  in  \cite{Vogel-2009, Bangert-2011}.   The  radiation
scheme used  within the  model to calculate  the vertical  profiles of
shortwave and longwave  radiative fluxes is GRAALS \cite{Ritter-1992}.
In order to account for  the interaction of aerosol particles with the
cloud microphysics and radiation,  COSMO-ART uses the two-moment cloud
microphysics  scheme  of Seifert  and  Beheng \cite{Seifert-2006}  and
comprehensive parameterizations for cloud condensation and ice crystal
nucleation  \cite{Bangert-2011, Bangert-2012}.   The  system of  stiff
ordinary differential equations described by chemical reactions in the
aqueous-phase together with the transfer reactions is solved using the
Kinetic PreProcessor (KPP, \cite{Damian-2002}).

Further model description details  can be found in \cite{Bangert-2012,
  Knote-2011, Knote-2013}.

### 2014-05-28 paper_enahpc14/Sec4:Experimental-results.tex
updates a few things according to the reviewers' suggestions

Figures~\ref{fig:1} and \ref{fig:2}  account respectively for \monch's
Isola  E1 Rack 2  and \pilat'  Isola HD  total power  measurements for
1-day or 2-day  simulations. On the Intel Ivy  Bridge EP based cluster
(i.e.   \monch), the  1-day simulation  was issued  only twice  due to
usage restrictions. As  time resolution was set to  one update every 5
minutes  for  power  sampling,  the  averaged  power  dissipation  was
computed by  considering 6 values for  each single run.   On the Intel
Xeon E5 based cluster (i.e.   \pilat), the 1-day simulation was issued
four times and  a 2-day run only once.   Similarly, the averaged power
dissipation was computed by considering 4 values for each single 1-day
run  and  9 values  for  the  2-days  run. Corresponding  results  are
gathered in Table~\ref{tab:1}.

@@

\begin{table}[htbf]
  \centering
  \caption{Averaged power dissipation (W) of the platforms.}
  \label{tab:1}
  \begin{tabular}{cccc}
    \hline\noalign{\smallskip}   \textbf{\scriptsize{Xeon   E5}}   &
    \textbf{\scriptsize{Ivy   Bridge}}   &  \textbf{\scriptsize{Xeon
        E5645}}   &   \textbf{\scriptsize{Xeon   E5645}}  \\   &   &
    \textbf{\scriptsize{(Polling)}}                                 &
    \textbf{\scriptsize{(Blocking)}}
    \\   \noalign{\smallskip}\hline\noalign{\smallskip}   18035.0  &
    12622.5 & 3713.6 & 3651.8 \\ \noalign{\smallskip}\hline
  \end{tabular}
\end{table}

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected,  Xeon  E5
outperforms Ivy  Bridge EP, being roughly 130\,\%  faster.  The reason
for that is two-fold: \emph{i}) it has higher clock frequency than Ivy
Bridge  (2.6\,GHz  against  2.2\,GHz),   and  \emph{ii})  it  aims  at
computing speed regardless to energy consumption.  In our experiments,
Ivy Bridge EP showed  the best energy-to-solution, reducing the energy
consumption of Xeon E5 by approximately 7\,\%.

@@

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    Xeon E5 and Ivy Bridge-EP architectures for a 24h simulation.}
  \label{fig:3}
\end{figure}

>>>>

Figures~\ref{fig:1} and \ref{fig:2}  account respectively for \monch's
Isola  E1 Rack 2  and \pilat'  Isola HD  total power  measurements for
1-day or 2-day simulations. On \monch, the 1-day simulation was issued
only twice due  to usage restrictions.  As time  resolution was set to
one  update every  5 minutes  for power  sampling, the  averaged power
dissipation was computed by considering  6 values for each single run.
On \pilat, the 1-day simulation was  issued four times and a 2-day run
only once.  Similarly, the  averaged power dissipation was computed by
considering 4  values for each single  1-day run and 9  values for the
2-days run. Corresponding results are gathered in Table~\ref{tab:1}.

@@

\begin{table}[htbf]
  \centering
  \caption{Averaged power dissipation (W) of the platforms.}
  \label{tab:1}
  \begin{tabular}{cccc}
    \hline\noalign{\smallskip}     \textbf{\scriptsize{PILATUS}}     &
    \textbf{\scriptsize{MONCH}}  &  \textbf{\scriptsize{TINTORRUM}}  &
    \textbf{\scriptsize{TINTORRUM}}           \\          &          &
    \textbf{\scriptsize{(Polling)}} & \textbf{\scriptsize{(Blocking)}}
    \\    \noalign{\smallskip}\hline\noalign{\smallskip}   18035.0   &
    12622.5 & 3713.6 & 3651.8 \\ \noalign{\smallskip}\hline
  \end{tabular}
\end{table}

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected,  Xeon  E5
outperforms Ivy  Bridge EP, being roughly 130\,\%  faster.  The reason
for that is two-fold: \emph{i}) it has higher clock frequency than Ivy
Bridge  (2.6\,GHz  against  2.2\,GHz),   and  \emph{ii})  it  aims  at
computing   speed  regardless   to  be   energy  efficient.    In  our
experiments,  Ivy  Bridge   EP  showed  the  best  energy-to-solution,
reducing the energy consumption of Xeon E5 by approximately 7\,\%.

@@

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    \pilat and \monch for a 24h simulation.}
  \label{fig:3}
\end{figure}
